A method for regularization of evolutionary polynomial regressions

### 2015-08-10 texts/submission_03_asc/eprr.tex
minor corrections

\begin{document}
%
%
\acrodef{GA}{Genetic algorithms}
\acrodef{rmse}[error]{root-mean-square}
\acrodef{EPR}[EPR]{Evolutionary Polynomial Regression}
\acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularization}
\acrodef{EPR}{Evolutionary Polynomial Regression}
\acrodef{EPRR}[EPRR]{Evolutionary Polynomial Regression with Regularization}
\acrodef{SVM}[SVM]{Support Vector Machines}
%
\setcounter{page}{0}
%
\begin{frontmatter}
%
\title{A method for regularization of evolutionary polynomial regressions}
%
%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}
%
\author{Francisco Coelho\corref{corfc}} 
\cortext[corfc]{Corresponding author: Tel.: +351-919-006-379}
\ead{fc@di.uevora.pt}
\address{Departamento de Inform\'{a}tica, Universidade de \'{E}vora, Rua Rom\~{a}o Ramalho 58, 7000-671 \'{E}vora, Portugal}
%
\author{Jo\~{a}o Pedro Neto}
\ead{jpn@di.fc.ul.pt}
\address{Departamento de Inform\'{a}tica, Faculdade de Ci\^{e}ncias da Universidade de Lisboa, Campo Grande, 1749-016 Lisboa, Portugal}
%
\begin{abstract}
While many applications require models that have no acceptable linear approximation, the simpler nonlinear models are defined by polynomials. The use of genetic algorithms to find polynomial models from data is known as Evolutionary Polynomial Regression.
%
This paper introduces Evolutionary Polynomial Regression with Regularization, an algorithm that extends the EPR method and describes a set of experiences on common datasets that compare both flavors of EPR and other methods including Linear Regression, Regression Trees and Support Vector Regression.

The empiric conclusion of those experiments is that EPR with regularization is able to achieve better fitting than other non-ensemble methods and it has shorter computation time than plain EPR.
\end{abstract}
%
\begin{keyword}
evolutionary polynomial regression \sep regularization \sep feature extraction
\end{keyword}
%\MSC 41A05\sep 41A10\sep 65D05\sep 65D17
%\KWD
%
\end{frontmatter}
%
%
\revised{ GAs are repeatedly mentioned as an approach to search for locally optimal parameters, this is simply incorrect and is a very surprising statement.  GAs are NOT a local optimizer in any shape or form.}{Our intention was merely to note that GAs give no guarantee of finding a global optimum}

\revised{What is the reference for the GA encoding used?}{It is ours.}

\revised{The penalty term introduces is the main contribution of this work, this should be clearly stated in the text. It is basically a complexity based penalty which is really not so novel, there are even several cases in literature of penalty based GA functions.  Please dig up further in the state of the art to back up better your claim of novelty.}{We included references to previous regularization in GA and outlined the role and originality of our contribution.}

@@

A simple exploration on the effect of the regularization parameter is depicted in Figure~\ref{fig:lambda.error-distribution} where it is possible to observe that penalizing polynomial complexity (i.e. $\lambda < 1$) achieves better results than the opposite ($\lambda > 1$). This observation motivates further inquire, done in section~\ref{sec:experimental.results}.\revised{please clarify this paragraph and justify this claim}{we rephrased the statement.}

@@

The performance of each method is evaluated on several common datasets. From each dataset 70\% of the observations are reserved for training purposes and the remaining observations used to estimate the \ac{rmse}. To enhance the robustness of results this process is repeated $25$~times, each time with a different shuffling of the samples in the train and test sets. Some datasets with attribute values of different magnitudes have a pre-processing scaling transformation. The box plots in figures \ref{artificial_dataset1_lambda1.0} and \ref{fig:four.datasets.summary} resume the test set error distributions over these different runs.

@@

\end{description}
%
\begin{figure*}[tb]\begin{center}
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{Fig3a.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig3b.pdf}
\\
\includegraphics[width=0.48\textwidth]{Fig3c.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig3d.pdf}
\end{tabular}
%
\caption{Graphical summary results for different regression methods on common datasets. Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.  The regression methods depicted in these figures are:
1. \ac{EPRR}, $\lambda = 0.7$;
2. \ac{EPRR}, $\lambda = 0.8$;
3. \ac{EPRR}, $\lambda = 0.9$;
4. \ac{EPR} (i.e. $\lambda = 1.0$);
5. Linear Regression;
6. \ac{SVM} (linear kernel);
7. Regression Trees;
8. Random Forest (with 100 trees);
9. Conditional Inference Trees.}
\label{fig:four.datasets.summary}
\label{Housing_dataset_lambda0.8_25runs}
\label{Abalone_dataset_lambda0.8_25runs}
\label{Auto-Mpg_dataset_lambda0.8_25runs}
\label{Kinematics300_lambda0.8_25runs}
\end{center}\end{figure*}

\subsection{Convergence speed}
%
\begin{figure*}[tb]
	\begin{center}
		\includegraphics[width=0.98\textwidth]{Fig4xx.pdf}
		\caption{Learning curve: Error progress for the \textsc{Housing} dataset during a single execution of the genetic algorithm. The figure shows the fitness evolution for different regularization values. The population of each run consists of $200$ polynomials.}
		\label{Abalone_fitnessProgress}
	\end{center}
\end{figure*}
%
Since this work is oriented to the \ac{rmse} of the \ac{EPRR} model it is necessary to assess how this depends on the number of generations of the \ac{GA}. As illustrated in Figure~\ref{Abalone_fitnessProgress}, the \ac{rmse} quickly drops during the initial $50$ to $100$ generations. Then, it proceeds slower achieving better solutions only with marginal error reduction.

@@

Of the regression methods considered \ac{SVM} achieves the best results in three out of four datasets. However \ac{SVM} and Conditional Inference Trees are pre-trained, having parameters tuned for each particular dataset unlike \ac{EPRR}, that runs with the same parameterization on all datasets. Even so it is the best estimator for the \textsc{Abalone} dataset and in the remaining datasets it outperforms most of the other estimators.

Comparing \ac{EPR} and \ac{EPRR} --- the main article's topic --- the regularized version achieves much better results at \textsc{Abalone} and especially \textsc{Kinematics}. On the \textsc{Housing} dataset errors are improved  wrt \ac{EPR} in a difference in means, resulted in a 95\% HDI (Highest Density Interval) equal to $\left\lbrack 0.001, 0.119 \right\rbrack$ which, while borderline, achieves statistical significance. Only in the \textsc{Auto MPG} dataset \ac{EPR} achieves better results, even if not that different from \ac{EPRR}.

@@

A cross-validation procedure can be implemented to refine the appropriate parameter values to achieve better errors. Namely, the regularization parameter, $\lambda$, can be tested with several values, instead of being fixed at $0.8$. Other parameters like mutation chance or the amount of elitism can also be tested. However, these type of tests need a low-level, fast implementation of \ac{EPR} and are postponed to future investigation.

@@

>>>>

\begin{document}
%
%
\acrodef{GA}{Genetic algorithms}
\acrodef{rmse}[error]{root-mean-square}
\acrodef{EPR}[EPR]{Evolutionary Polynomial Regression}
\acrodef{EPRR}[EPR$^2$]{Evolutionary Polynomial Regression with Regularization}
\acrodef{EPR}{Evolutionary Polynomial Regression}
\acrodef{EPRR}[EPRR]{Evolutionary Polynomial Regression with Regularization}
\acrodef{SVM}[SVM]{Support Vector Machines}
\acrodef{RF}[RF]{Random Forests}

%
\setcounter{page}{0}
%
\begin{frontmatter}
%
\title{A method for regularization of evolutionary polynomial regressions}
%
%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}
%
\author{Francisco Coelho\corref{corfc}} 
\cortext[corfc]{Corresponding author: Tel.: +351-919-006-379}
\ead{fc@di.uevora.pt}
\address{Departamento de Inform\'{a}tica, Universidade de \'{E}vora, Rua Rom\~{a}o Ramalho 58, 7000-671 \'{E}vora, Portugal}
%
\author{Jo\~{a}o Pedro Neto}
\ead{jpn@di.fc.ul.pt}
\address{University of Lisboa, Faculty of Sciences, BioISI– Biosystems \& Integrative Sciences Institute, Campo Grande, 1749-016 Lisboa, Portugal}
%
\begin{abstract}
While many applications require models that have no acceptable linear approximation, the simpler nonlinear models are defined by polynomials. The use of genetic algorithms to find polynomial models from data is known as Evolutionary Polynomial Regression.
%
This paper introduces Evolutionary Polynomial Regression with Regularization, an algorithm that extends the EPR method with a regularization term to control polynomial complexity. The article also describes a set of experiences on common datasets that compare both flavors of EPR and other methods including Linear Regression, Regression Trees and Support Vector Regression.

The empiric conclusion of those experiments is that EPR with regularization is able to achieve better fitting than other non-ensemble methods and it needs less computation time than plain EPR.
\end{abstract}
%
\begin{keyword}
evolutionary polynomial regression \sep regularization \sep feature extraction
\end{keyword}
%\MSC 41A05\sep 41A10\sep 65D05\sep 65D17
%\KWD
%
\end{frontmatter}
%
%
\revised{ GAs are repeatedly mentioned as an approach to search for locally optimal parameters, this is simply incorrect and is a very surprising statement.  GAs are NOT a local optimizer in any shape or form.}{Our intention was merely to note that GAs give no guarantee of finding a global optimum. We rephrase the mentioned sentences.}

\revised{What is the reference for the GA encoding used?}{It is ours. We stated this contribution more explicitly.}

\revised{The penalty term introduces is the main contribution of this work, this should be clearly stated in the text. It is basically a complexity based penalty which is really not so novel, there are even several cases in literature of penalty based GA functions.  Please dig up further in the state of the art to back up better your claim of novelty.}{We included references to previous regularization in GA and better outlined the role and originality of our contribution.}

@@

A simple exploration on the effect of the regularization parameter is depicted in Figure~\ref{fig:lambda.error-distribution} where it is possible to observe that penalizing polynomial complexity (i.e. $\lambda < 1$) achieves better results than the opposite ($\lambda > 1$). This observation motivates further inquire, done in section~\ref{sec:experimental.results} where we compare the regularized versions with $\lambda < 1$ and the non-regularized algorithm against several other regression methods.\revised{please clarify this paragraph and justify this claim}{we rephrased the statement.}

@@

The performance of each method is evaluated on several common datasets. From each dataset 70\% of the observations are reserved for training purposes and the remaining observations used to estimate the \ac{rmse}. To enhance the robustness of results this process is repeated $75$~times, each time with a different shuffling of the samples in the train and test sets. The datasets have been scaled to prevent problems with different magnitude attributes. The box plots in figures \ref{artificial_dataset1_lambda1.0} and \ref{fig:four.datasets.summary} resume the test set error distributions over these different runs.

@@

\end{description}

%
\begin{figure*}[tb]\begin{center}
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{Fig3a.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig3b.pdf}
\\
\includegraphics[width=0.48\textwidth]{Fig3c.pdf}
&
\includegraphics[width=0.48\textwidth]{Fig3d.pdf}
\end{tabular}
%
\caption{Graphical summary results for different regression methods on common datasets. Although \ac{EPRR} not always achieves the smallest expected error, performance is on-par with more sophisticated methods.  The regression methods depicted in these figures are:
1. \ac{EPRR}, $\lambda = 0.7$;
2. \ac{EPRR}, $\lambda = 0.8$;
3. \ac{EPRR}, $\lambda = 0.9$;
4. \ac{EPR} (i.e. $\lambda = 1.0$);
5. Linear Regression;
6. \ac{SVM} (linear kernel);
7. Regression Trees;
8. Random Forest (with 100 trees);
9. Conditional Inference Trees.}
\label{fig:four.datasets.summary}
\label{Housing_dataset_lambda0.8_25runs}
\label{Abalone_dataset_lambda0.8_25runs}
\label{Auto-Mpg_dataset_lambda0.8_25runs}
\label{Kinematics300_lambda0.8_25runs}
\end{center}\end{figure*}

\subsection{Convergence speed}
%
\begin{figure*}[tb]
	\begin{center}
		\includegraphics[width=0.98\textwidth]{Fig4xx.pdf}
		\caption{Learning curve: Error progress for the \textsc{Housing} dataset during a single execution of the genetic algorithm. The figure shows the fitness evolution for different regularization values. The population of each run consists of $200$ polynomials.}
		\label{Abalone_fitnessProgress}
	\end{center}
\end{figure*}
%
Since this work is oriented to the \ac{EPRR} model \ac{rmse} it is necessary to assess how this depends on the number of generations of the \ac{GA}. As illustrated in Figure~\ref{Abalone_fitnessProgress}, the \ac{rmse} quickly drops during the initial $100$ generations. Then, it proceeds slower achieving better solutions only with marginal error reduction. The \ac{EPR} takes more than $300$ generations to achieve similar error results.

@@

Of the regression methods considered \ac{RF} achieves the best results in two out of four datasets. If we compare against non-ensemble methods, \ac{EPR} and \ac{EPRR} outperform the other estimators, with the exception of Regression Trees in the Kinematics dataset.

Comparing \ac{EPR} and \ac{EPRR} --- the main article's topic --- the regularized version achieves better results at \textsc{Abalone} and especially \textsc{Kinematics}. A Bayesian Estimation was computed, which estimates the difference in means between two groups and yields a probability distribution over the difference. On the Abalone dataset errors are improved wrt EPR in a difference in means, for all tested $\lambda$ values less than $1$, resulted in a 95\% HDI (Highest Density Interval) equal to $\left\lbrack 0.01, 0.02 \right\rbrack$ which achieves statistical significance. On the Auto-MPG dataset the 95\% HDI is $\left\lbrack -0.01, 0.008 \right\rbrack$ which includes zero. On the Housing dataset the 95\% HDI is $\left\lbrack- 0.02, 0.05 \right\rbrack$, for $\lambda = 0.7$, which includes zero (the other values of $\lambda$ had similar statistics).

@@

Parameters like mutation chance or the amount of elitism can be tested for tuning purposes. However, these type of tests need a low-level, fast implementation of \ac{EPR} and are postponed to future investigation.

@@

Jo\~ao Pedro Neto's work is supported by centre grant (to BioISI, Centre Reference: UID/MULTI/04046/2013), from FCT/MCTES/PIDDAC, Portugal.
